Trajectories

### 2021-09-30 Rmd/report.Rmd
Merge into one manuscript
Fix comprehension model and update report

We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age` by `item`, and their correlation parameter. We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope. Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`. The models were implemented in `brms` as:

* `understands/produces ~ 1 + age + frequency + (1 + age | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age + frequency + doe + (1 + age + doe | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age + frequency + doe*cognate + (1 + age + doe | item) + (1 + frequency + cognate| id)`

@@

```
understands ~
1 + age + frequency + lp*cognate +
(1 + age + lp | te) +
(1 + frequency + cognate | id),
family =  bernoulli("logit")
```

@@

```{stan stan_model, echo=FALSE, output.var="model"}

@@

## Prior-predictive checks (PPC)

### PPC: Comprehension

```{r ppc_comp, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 30),
  doe_center = range(responses$doe_center),
  cognate = unique(items$cognate),
  frequency_center = 0
)

m <- add_fitted_draws(nd, fits_comp_prior$fit_2, n = 20, re_formula = NA) %>% 
  mutate(
    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
  )

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_grid(~doe_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Comprehension)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

### PPC: Production

```{r ppc_prod, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3, fig.width=6}

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 30),
  doe_center = range(responses$doe_center),
  cognate = unique(items$cognate),
  frequency_center = 0
)

m <- add_fitted_draws(nd, fits_prod_prior$fit_2, n = 20, re_formula = NA) %>% 
  mutate(
    frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")),
    doe_center = factor(doe_center, levels = unique(nd$doe_center), labels = c("DoE: 0%", "DoE: 100%"))
  )

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_grid(~doe_center) +
  geom_line(aes(group = interaction(.draw,  frequency_center, cognate)), alpha = 0.5) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.05) +
  # stat_summary(fun.data = "mean_se", geom = "pointrange") +
  # stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "P(Production)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits_comp$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  ) %>% 
  as_raw_html()

@@

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 30),
  doe_center = range(responses$doe_center),
  cognate = unique(items$cognate),
  frequency_center = 0
)

@@

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_wrap(~doe_center) +
  geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.25) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  stat_summary(fun = "mean", geom = "line", size = 1) +  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "Posterior probability of comprehension\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

ggplot(m, aes(age, .value, colour = cognate, fill = cognate)) +
  facet_wrap(~doe_center) +
  geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.1, size = 1) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "Posterior probability of production\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

>>>>

We initially fitted a null model (`fit_0`) than only included the predictors `age` and `frequency` as nuisance parameters, along with random intercepts by `id` and `item`, and random slopes of `frequency` by `id`, and `age_center` by `item`, and their correlation parameter. We then expanded this model (`fit_1`) to include the main effect of `doe`, and the `doe` by `item` random slope. Finally, we added the main effect `cognate` (`fit_2`), its interaction with `doe` (`doe:cognate`), and  random slopes for `cognate` by `id`. The models were implemented in `brms` as:

* `understands/produces ~ 1 + age_center + frequency + (1 + age_center | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age_center + frequency + doe + (1 + age_center + doe | item) + (1 + frequency | id)`
* `understands/produces ~ 1 + age_center + frequency + doe*cognate + (1 + age_center + doe | item) + (1 + frequency + cognate| id)`

@@

```
understands ~
1 + age_center + frequency + lp*cognate +
(1 + age_center + lp | te) +
(1 + frequency + cognate | id),
family =  bernoulli("logit")
```

@@

@@

@@

```{r fixed_effects_table, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fits_comp$fit_2)$fixed %>% 
  as.data.frame() %>%
  rownames_to_column("term") %>% 
  mutate(
    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
  ) %>% 
  gt() %>% 
  fmt_percent(2:5) %>% 
  fmt_number(6:6, decimals = 2) %>% 
  fmt_number(7:8, decimals = 0) %>% 
  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
  cols_label(
    term = md("**Predictor**"),
    Estimate = md("**Mean**"),
    Est.Error = md("**SEM**"),
    "l-95% CI" = md("**95\\% CrI**"),
    Rhat = md("**Rhat**"),
    Bulk_ESS = md("**Bulk ESS**"),
    Tail_ESS = md("**Tail ESS**")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the inverse logit to get the average probability of correct response",
    locations = cells_body(columns = "term", rows = term=="Intercept")
  ) %>% 
  tab_footnote(
    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
    locations = cells_body(columns = "term", rows = term %in% c("age_center", "frequency", "doe_center", "cognate1", "doe_center:cognate1"))
  ) %>% 
  tab_footnote(
    footnote = "ESS: Effective sample size",
    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
  ) %>% 
  cols_align(
    align = c("center"),
    columns = 2:4
  ) %>% 
  as_raw_html()

@@

nd <- expand.grid(
  age_center = seq(min(responses$age_center), max(responses$age_center), length.out = 30),
  doe_center = range(responses$doe_center),
  cognate = unique(items$cognate),
  frequency_center = 0
)

@@

ggplot(m, aes(age_center, .value, colour = cognate, fill = cognate)) +
  facet_wrap(~doe_center) +
  geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.25) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  stat_summary(fun = "mean", geom = "line", size = 1) +  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "Posterior probability of comprehension\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )

@@

ggplot(m, aes(age_center, .value, colour = cognate, fill = cognate)) +
  facet_wrap(~doe_center) +
  geom_line(aes(group = interaction(.draw, cognate), colour = cognate), alpha = 0.1, size = 1) +
  # stat_lineribbon(size = 0, alpha = 0.5, .width = 0.95) +
  stat_summary(fun = "mean", geom = "line", size = 1) +
  geom_hline(yintercept = 0.5, size = 1, colour = "grey") +
  labs(x = "Age (months)", y = "Posterior probability of production\n(Linear prediction)", colour = "Cognateness", fill = "Cognateness") +
  scale_y_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    legend.title = element_blank()
  )
